---
title: "DS6306_CaseStudy02"
author: "Rudranil Mitra"
date: "8/14/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(caret)
library(class)
library(dplyr)
library(e1071)
library(FNN) 
library(gmodels) 
library(psych)
library(epiR)
library(DMwR)
```


```{r}
# Load attrition dataset with labels
attrition_dataset = read.csv('D:\\SMU_MSDS\\MSDS_6306_Doing-Data-Science\\Unit 14 and 15 Case Study 2\\CaseStudy2-data.csv')
# Load attrition dataset without labels 
attrition_dataset_wl = read.csv('D:\\SMU_MSDS\\MSDS_6306_Doing-Data-Science\\Unit 14 and 15 Case Study 2\\CaseStudy2CompSet No Attrition.csv')
# 
attrition_dataset_lm = attrition_dataset
```

### Does mean age differ significantly among Attrition Groups 
```{r}
# Execute the t-test between two groups, we are assuming that the variances are not equal
fit = t.test(
  attrition_dataset[attrition_dataset['Attrition']=='No',]$Age,
  attrition_dataset[attrition_dataset['Attrition']=='Yes',]$Age,
  var.equal = FALSE)
# P-value is less than the significance level(0.05), so we can reject the null hypothesis and say that mean age is different between two groups
fit$p.value
# Looking at the confidence intervals we can suggest that individuals who are  not leaving the company have higher mean age than the individuals who leave the company
fit$conf.int
attrition_dataset %>% ggplot(mapping=aes(x=Age,y=Attrition,fill=Attrition)) + 
  geom_boxplot() + 
  annotate("text",x=40,y=2.5,label=paste0('P-value=',round(fit$p.value,4))) + 
  ggtitle('Distribution of Age between two Attrition Groups')

```
### Does median monthly income differ significantly among Attrition Groups 
```{r}
# Execute the t-test between two groups, we are assuming that the variances are not equal
fit = t.test(
  log(attrition_dataset[attrition_dataset['Attrition']=='No',]$MonthlyIncome),
  log(attrition_dataset[attrition_dataset['Attrition']=='Yes',]$MonthlyIncome),
  var.equal = FALSE)
# P-value is less than the significance level(0.05), so we can reject the null hypothesis and say that median monthly income is different between two groups
fit$p.value
# Looking at the confidence intervals(confidence intervals are in the log scale) we can suggest that individuals who are  not leaving the company have higher median monthly income than the individuals who leave the company. So monthly income is a significant reason for leaving the company
# The lower bound of the increase between two groups is 27% 
round(exp(fit$conf.int[1])-1,2)
# The upper bound of the increase between two groups is 60%
round(exp(fit$conf.int[2])-1,2)
attrition_dataset %>% ggplot(mapping=aes(x=MonthlyIncome,y=Attrition,fill=Attrition)) + 
  geom_boxplot() + 
  annotate("text",x=15000,y=2.5,label=paste0('P-value=',round(fit$p.value,4))) + 
  ggtitle('Distribution of Monthly Income between two Attrition Groups')


```


### Does mean distance from home differ significantly among Attrition Groups 
```{r}
# Execute the t-test between two groups, we are assuming that the variances are not equal
fit = t.test(
  attrition_dataset[attrition_dataset['Attrition']=='No',]$DistanceFromHome,
  attrition_dataset[attrition_dataset['Attrition']=='Yes',]$DistanceFromHome,
  var.equal = FALSE)
# P-value is less than the significance level(0.05), so we can reject the null hypothesis and say that mean distance from home is different between two groups
fit$p.value
# Looking at the confidence intervals we can suggest that individuals who are not leaving the company are staying close to home than the individuals who are leaving the company
fit$conf.int
attrition_dataset %>% ggplot(mapping=aes(x=DistanceFromHome,y=Attrition,fill=Attrition)) + 
  geom_boxplot() + 
  annotate("text",x=10,y=2.5,label=paste0('P-value=',round(fit$p.value,4))) + 
  ggtitle('Distribution of Distance From Home between two Attrition Groups')

```
## Does mean total working years differ significantly among Attrition Groups 
```{r}
# Execute the t-test between two groups, we are assuming that the variances are not equal
fit = t.test(
  attrition_dataset[attrition_dataset['Attrition']=='No',]$TotalWorkingYears,
  attrition_dataset[attrition_dataset['Attrition']=='Yes',]$TotalWorkingYears,
  var.equal = FALSE)
# P-value is less than the significance level(0.05), so we can reject the null hypothesis and say that mean total working years is different between two groups
fit$p.value
attrition_dataset %>% ggplot(mapping=aes(x=DistanceFromHome,y=Attrition,fill=Attrition)) + 
  geom_boxplot() + 
    annotate("text",x=20,y=2.5,label=paste0('P-value=',round(fit$p.value,4))) + 
  ggtitle('Distribution of Total Working Years between two Attrition Groups')

```

## Does mean total years working at the company different significantly among Attrition Groups
```{r}
# Execute the t-test between two groups, we are assuming that the variances are not equal
fit = t.test(
  attrition_dataset[attrition_dataset['Attrition']=='No',]$YearsAtCompany,
  attrition_dataset[attrition_dataset['Attrition']=='Yes',]$YearsAtCompany,
  var.equal = FALSE)
# P-value is less than the significance level(0.05), so we can reject the null hypothesis and say that mean years at current company is different between two groups
fit$p.value
attrition_dataset %>% ggplot(mapping=aes(x=DistanceFromHome,y=Attrition,fill=Attrition)) + 
  geom_boxplot() + 
  annotate("text",x=20,y=2.5,label=paste0('P-value=',round(fit$p.value,4))) + 
  ggtitle('Distribution of Years at Current company between two Attrition Groups')

```
## Run categorical tests to check whether those variables are associated with attrition
```{r}
fit=chisq.test(table(attrition_dataset$BusinessTravel,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$BusinessTravel),
        col = c("green","red"),
        main=paste0('Attrition by Business Travel, p-value=',round(fit$p.value,4)),
        xlab='Business Travel Class') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))


fit=chisq.test(table(attrition_dataset$Department,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$Department),
        col = c("green","red"),
        main=paste0('Attrition by Department, p-value=',round(fit$p.value,4)),
        xlab='Department') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))

fit=chisq.test(table(attrition_dataset$EducationField,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$EducationField),
        col = c("green","red"),
        main=paste0('Attrition by Education Field, p-value=',round(fit$p.value,4)),
        xlab='Education Field') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))

fit=chisq.test(table(attrition_dataset$Gender,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$Gender),
        col = c("green","red"),
        main=paste0('Attrition by Gender, p-value=',round(fit$p.value,4)),
        xlab='Gender') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))

fit=chisq.test(table(attrition_dataset$JobRole,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$JobRole),
        col = c("green","red"),
        main=paste0('Attrition by JobRole, p-value=',round(fit$p.value,4)),
        xlab='Job Role') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))

fit=chisq.test(table(attrition_dataset$MaritalStatus,attrition_dataset$Attrition))
barplot(table(attrition_dataset$Attrition,attrition_dataset$MaritalStatus),
        col = c("green","red"),
        main=paste0('Attrition by Marital Status, p-value=',round(fit$p.value,4)),
        xlab='Marital Status') 
legend("topleft",c("Attrition - Yes","Attrition - No"),fill = c("red","green"))
```

## Convert categorical variables by using dummy encoding 
```{r}
clean_wrangle_dataset = function(df,test){
  ##
  input_dataset = df
  ##
  if(test==0){
    input_dataset$Attrition = as.factor(input_dataset$Attrition)  
  }
  ##
  input_dataset$BusinessTravel = as.factor(input_dataset$BusinessTravel)
  input_dataset$Department = as.factor(input_dataset$Department)
  input_dataset$EducationField = as.factor(input_dataset$EducationField)
  input_dataset$Gender = as.factor(input_dataset$Gender)
  input_dataset$JobRole = as.factor(input_dataset$JobRole)
  input_dataset$MaritalStatus = as.factor(input_dataset$MaritalStatus)
  input_dataset$Over18 = as.factor(input_dataset$Over18)
  input_dataset$OverTime = as.factor(input_dataset$OverTime)
  
  # Shorten Department Names by replacing the values 
  
  input_dataset$Department = str_replace(input_dataset$Department, "Research & Development", "RnD")
  input_dataset$Department = str_replace(input_dataset$Department, "Human Resources", "HR")
  
  
  # Change variables with 2 levels to 1 and 0 
  input_dataset$Gender <- ifelse(input_dataset$Gender == "Female", 1, 0)
  input_dataset$OverTime <- ifelse(input_dataset$OverTime == "Yes", 1, 0)
  
  # Remove Variables not needed in the model
  input_dataset = input_dataset %>% select(-Over18)
  input_dataset = input_dataset %>% select(-EmployeeCount)
  input_dataset = input_dataset %>% select(-EmployeeNumber)
  input_dataset = input_dataset %>% select(-ID)
  input_dataset = input_dataset %>% select(-StandardHours)
  input_dataset = input_dataset %>% select(-DailyRate)
  input_dataset = input_dataset %>% select(-HourlyRate)
  input_dataset = input_dataset %>% select(-MonthlyRate)
  
  
  
  # Scale numeric variables 
  input_dataset[, c("Age","DistanceFromHome","Education","EnvironmentSatisfaction","JobInvolvement","JobLevel","JobSatisfaction","MonthlyIncome","NumCompaniesWorked","PercentSalaryHike","PerformanceRating","RelationshipSatisfaction","StockOptionLevel","TotalWorkingYears","TrainingTimesLastYear","WorkLifeBalance","YearsAtCompany","YearsInCurrentRole","YearsSinceLastPromotion","YearsWithCurrManager")] =  scale(input_dataset[, c("Age","DistanceFromHome","Education","EnvironmentSatisfaction","JobInvolvement","JobLevel","JobSatisfaction","MonthlyIncome","NumCompaniesWorked","PercentSalaryHike","PerformanceRating","RelationshipSatisfaction","StockOptionLevel","TotalWorkingYears","TrainingTimesLastYear","WorkLifeBalance","YearsAtCompany","YearsInCurrentRole","YearsSinceLastPromotion","YearsWithCurrManager")])
  
  
  # Add feature names and replace spaces
  
  input_dataset$BusinessTravel = paste("BT_",str_replace(input_dataset$BusinessTravel," ","_"),sep="")
  input_dataset$Department = paste("DP_",str_replace(input_dataset$Department," ","_"),sep="")
  input_dataset$EducationField = paste("EF_",str_replace(input_dataset$EducationField," ","_"),sep="")
  input_dataset$JobRole = paste("JR_",str_replace(input_dataset$JobRole," ","_"),sep="")
  input_dataset$MaritalStatus =  paste("MS_",str_replace(input_dataset$MaritalStatus," ","_"),sep="")
  
  
  # Dummy code categorical variables having 3 or more levels
  BusinessTravel = as.data.frame(dummy.code(input_dataset$BusinessTravel))
  Department = as.data.frame(dummy.code(input_dataset$Department))      
  EducationField = as.data.frame(dummy.code(input_dataset$EducationField))
  JobRole = as.data.frame(dummy.code(input_dataset$JobRole))
  MaritalStatus = as.data.frame(dummy.code(input_dataset$MaritalStatus))
  
  
  # Add the dummy codes to the dataset
  input_dataset = cbind(input_dataset,BusinessTravel,Department,EducationField,JobRole,MaritalStatus)
  
  # Remove the original categorical variables 
  input_dataset = input_dataset %>% select(-one_of(c("BusinessTravel", "Department", "EducationField","JobRole","MaritalStatus")))
  
  # Return the dataset
  return(input_dataset)
}
```

## Clean and make the datasets ready for KNN algorithm
```{r}
attrition_dataset = clean_wrangle_dataset(attrition_dataset,0)
#Move outcome to the last column
attrition_dataset$Outcome = attrition_dataset$Attrition
attrition_dataset = attrition_dataset %>% select(-Attrition)

attrition_dataset_wl = clean_wrangle_dataset(attrition_dataset_wl,1)


```

## Run KNN on training set to check specificity and sensitivity of the model 
```{r}
#set.seed(1243) # set the seed to make the partition reproducible

# 80% of the sample size
smp_size <- floor(0.8 * nrow(attrition_dataset))
train_ind <- sample(seq_len(nrow(attrition_dataset)), size = smp_size)

train_df <- attrition_dataset[train_ind, ]
test_df <- attrition_dataset[-train_ind, ]
# Use SMOTE to oversample the Attrition = Yes Observations

train_df = SMOTE(Outcome~.,train_df,perc.over = 600,perc.under=100,k=10)
prop.table(table(train_df$Outcome))

classifications = knn(train_df[,c(1:46)], test_df[,c(1:46)], train_df$Outcome,prob = T,k=10)
confusionMatrix(table(test_df$Outcome,classifications,dnn=c('Predicted','Actual')))
```
## Predict classifications on the test data set 
```{r}
# attrition_dataset_train = SMOTE(Outcome~.,attrition_dataset,perc.over = 600,perc.under=100,k=10)
# classifications = knn(attrition_dataset_train[,c(1:46)], attrition_dataset_wl[,c(1:46)], attrition_dataset_train$Outcome,prob = F,k=5)
# write.csv(x=classifications, file='D:\\SMU_MSDS\\MSDS_6306_Doing-Data-Science\\Unit 14 and 15 Case Study 2\\attrition_results.csv',row.names = F)
```
## Run linear regression
```{r}
attrition_dataset_lm$BusinessTravel = as.factor(attrition_dataset_lm$BusinessTravel)
attrition_dataset_lm$Department =as.factor(attrition_dataset_lm$Department)
attrition_dataset_lm$Gender = as.factor(attrition_dataset_lm$Gender)
attrition_dataset_lm$JobRole = as.factor(attrition_dataset_lm$JobRole)
attrition_dataset_lm$OverTime = as.factor(attrition_dataset_lm$OverTime)
      
model.full = lm(MonthlyIncome~Age+
                BusinessTravel+
                Department+
                Education+
                EducationField+
                EnvironmentSatisfaction+
                Gender+
                JobInvolvement+
                JobLevel+
                JobRole+
                JobSatisfaction+
                NumCompaniesWorked+
                OverTime+
                PercentSalaryHike+
                PerformanceRating+
                RelationshipSatisfaction+
                StockOptionLevel+
                TotalWorkingYears+
                TrainingTimesLastYear+
                WorkLifeBalance+
                YearsAtCompany+
                YearsInCurrentRole+
                YearsSinceLastPromotion+
                YearsWithCurrManager,data=attrition_dataset_lm)
      
model.aic.backward <- step(model.full, direction = "backward", trace = 1)

fit=lm(MonthlyIncome ~ BusinessTravel + Gender + JobLevel + JobRole + 
  PercentSalaryHike + PerformanceRating + TotalWorkingYears + 
  YearsSinceLastPromotion + YearsWithCurrManager,data=attrition_dataset_lm)
summary(fit)
par(mfrow=c(2,2))
plot(fit)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
